---
title: "Model summaries: Viral disease outcomes are indistinguishable between experimentally infected bats and rodents"
author: "Maxwell J. Farrell"
# date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
  html_document:
    highlight: default
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: true
    collapsed: true
    df-print: paged
    theme: spacelab
urlcolor: blue

---



```{r loading packages, echo=F, message=F, warning=F}
# Packages 
require(ape)#;packageVersion("ape")#5.5
require(dplyr)#;packageVersion("dplyr")#1.1.4
require(tidyr)#;packageVersion("tidyr")#1.3.0
require(ggplot2)#;packageVersion("ggplot2")#3.4.4
require(ggtree)#;packageVersion("ggtree")#3.0.4
require(lubridate)#;packageVersion("lubridate")#1.8.0
require(cowplot)#;packageVersion("cowplot")#1.1.1
require(parameters)#;packageVersion("parameters")#0.21.6
require(knitr)#;packageVersion("knitr")#1.4.9
require(kableExtra)#;packageVersion("kableExtra")#1.3.4
require(tidybayes)#;packageVersion("tidybayes")#3.0.6
require(brms)#;packageVersion("brms")#2.21.6
require(bayesplot)#;packageVersion("bayesplot")#1.8.1
bayesplot_theme_set(theme_bw())

options(dplyr.summarise.inform = FALSE)

# set colours for bats and rodents
colours_BR <- c("#AB1808","#1B85BF")

# set seed
set.seed(1989666)

# markdown setup
knitr::opts_chunk$set(echo=T, eval=T, message=F, warning=F) 

# plotting functions
source('./scaling_plotting_functions.R')

```


# Full models

```{r plots_full_models, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_full.rds")
s_mort <- readRDS("../fit_models/spec_mort_full.rds")
s_sev <- readRDS("../fit_models/spec_severity_full.rds")

i_dis <- readRDS("../fit_models/indiv_disease_full.rds")
i_mort <- readRDS("../fit_models/indiv_mort_full.rds")
i_sev <- readRDS("../fit_models/indiv_severity_full.rds")

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals.pdf", cp, width=11.5, height=3.5)

```


# Reservoir Match Models

## Adding Reservoir Match

```{r plots_resMatch_models, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_resMatch.rds")
s_mort <- readRDS("../fit_models/spec_mort_resMatch.rds")
s_sev <- readRDS("../fit_models/spec_severity_resMatch.rds")

i_dis <- readRDS("../fit_models/indiv_disease_resMatch.rds")
i_mort <- readRDS("../fit_models/indiv_mortality_resMatch.rds")
i_sev <- readRDS("../fit_models/indiv_severity_resMatch.rds")

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals_resMatch.pdf", cp, width=11.5, height=3.5)


```

## Reservoir Match without Host Evolutionary Isolation

```{r plots_resMatch_noHEI_models, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_resMatch_alone.rds")
s_mort <- readRDS("../fit_models/spec_mort_resMatch_alone.rds")
s_sev <- readRDS("../fit_models/spec_severity_resMatch_alone.rds")

i_dis <- readRDS("../fit_models/indiv_disease_resMatch_alone.rds")
i_mort <- readRDS("../fit_models/indiv_mortality_resMatch_alone.rds")
i_sev <- readRDS("../fit_models/indiv_severity_resMatch_alone.rds")

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals_resMatch_alone.pdf", cp, width=11.5, height=3.5)


```

## Reservoir Match x Host Order Interaction

```{r plots_resMatch_inter, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_resMatch_inter.rds")
s_mort <- readRDS("../fit_models/spec_mortality_resMatch_inter.rds")
s_sev <- readRDS("../fit_models/spec_severity_resMatch_inter.rds")

i_dis <- readRDS("../fit_models/indiv_disease_resMatch_inter.rds")
i_mort <- readRDS("../fit_models/indiv_mortality_resMatch_inter.rds")
i_sev <- readRDS("../fit_models/indiv_severity_resMatch_inter.rds")

## Joint plot

padding <- theme(plot.margin = unit(c(t=0.1, r=0.1, b=0, l=0), "cm"))

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals_resMatch_inter.pdf", cp, width=11.5, height=3.5)

# padding <- theme(plot.margin = unit(c(t=0.1, r=0.1, b=0, l=0), "cm"))
    
# cp <- egg::ggarrange(        
#     (intervals_plot(i_dis) + padding + theme(axis.text.y=element_text(size=12)) + ggtitle("A) Disease presence")), 
#     (intervals_plot(i_mort) + padding + theme(axis.text.y=element_blank()) + ggtitle("B) Mortality")), 
#     (intervals_plot(i_sev) + padding + theme(axis.text.y=element_blank(), legend.position = "right") + ggtitle("C) Severity")), ncol=3, padding=0.5)

# ggsave("../plots_tables/combined_individual_intervals_resMatch_inter.pdf", cp, width=11.5, height=3.5)

cp <- egg::ggarrange(        
    (plot_conditional_effects(s_dis, "Probability of disease outcome") + padding + theme(axis.text.y=element_text(size=12), legend.position="none") + ggtitle("A) Disease presence")), 
    (plot_conditional_effects(s_mort, "") + padding + theme(axis.text.y=element_blank(), legend.position="none") + ggtitle("B) Mortality")), 
    (plot_conditional_effects(s_sev, "") + padding + theme(axis.text.y=element_blank(), legend.position = "right") + ggtitle("C) Severity")), ncol=3, padding=0.5)

ggsave("../plots_tables/species_conditional_effects_resMatch_inter.pdf", cp, width=11.5, height=3.5)

cp2 <- egg::ggarrange(        
    (plot_conditional_effects(i_dis, "Probability of disease outcome") + padding + theme(axis.text.y=element_text(size=12), legend.position="none") + ggtitle("A) Disease presence")), 
    (plot_conditional_effects(i_mort, "") + padding + theme(axis.text.y=element_blank(), legend.position="none") + ggtitle("B) Mortality")), 
    (plot_conditional_effects(i_sev, "") + padding + theme(axis.text.y=element_blank(), legend.position = "right") + ggtitle("C) Severity")), ncol=3, padding=0.5)


ggsave("../plots_tables/individual_conditional_effects_resMatch_inter.pdf", cp2, width=11.5, height=3.5)


```



# Removing Lyssaviruses

## Full model excluding lyssaviruses

```{r plots_nolys_models, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_nolys.rds")
s_mort <- readRDS("../fit_models/spec_mort_nolys.rds")
s_sev <- readRDS("../fit_models/spec_severity_nolys.rds")

i_dis <- readRDS("../fit_models/indiv_disease_nolys.rds")
i_mort <- readRDS("../fit_models/indiv_mort_nolys.rds")
i_sev <- readRDS("../fit_models/indiv_severity_nolys.rds")

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals_nolys.pdf", cp, width=11.5, height=3.5)


```

## No lyssaviruses with reservoir match added

```{r plots_nolys_resMatch_models, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_resMatch_nolys.rds")
s_mort <- readRDS("../fit_models/spec_mort_resMatch_nolys.rds")
s_sev <- readRDS("../fit_models/spec_severity_resMatch_nolys.rds")

i_dis <- readRDS("../fit_models/indiv_disease_resMatch_nolys.rds")
i_mort <- readRDS("../fit_models/indiv_mortality_resMatch_nolys.rds")
i_sev <- readRDS("../fit_models/indiv_severity_resMatch_nolys.rds")

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals_resMatch_nolys.pdf", cp, width=11.5, height=3.5)


```


# Restricting to Homologous and Heterologous inoculations

## Homologous inoculations

```{r plots_homoHosts_models, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_homoHosts.rds")
s_mort <- readRDS("../fit_models/spec_mort_homoHosts.rds")
s_sev <- readRDS("../fit_models/spec_severity_homoHosts.rds")

i_dis <- readRDS("../fit_models/indiv_disease_homoHosts.rds")
i_mort <- readRDS("../fit_models/indiv_mort_homoHosts.rds")
i_sev <- readRDS("../fit_models/indiv_severity_homoHosts.rds")

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals_homoHosts.pdf", cp, width=11.5, height=3.5)


```


## Heterologous inoculations

```{r plots_heteroHosts_models, fig.width=10}

s_dis <- readRDS("../fit_models/spec_disease_heteroHosts.rds")
s_mort <- readRDS("../fit_models/spec_mort_heteroHosts.rds")
s_sev <- readRDS("../fit_models/spec_severity_heteroHosts.rds")

i_dis <- readRDS("../fit_models/indiv_disease_heteroHosts.rds")
i_mort <- readRDS("../fit_models/indiv_mort_heteroHosts.rds")
i_sev <- readRDS("../fit_models/indiv_severity_heteroHosts.rds")

cp <- six_panel_forest_plot(s_dis, i_dis, s_mort, i_mort, s_sev, i_sev)

ggsave("../plots_tables/combined_species_individual_intervals_heteroHosts.pdf", cp, width=11.5, height=3.5)


```


# Alternative dose adjustments

## Raw dose (no body mass adjustment)

```{r plots_dose_amount, fig.width=10}
  
i_dis <- readRDS("../fit_models/indiv_disease_dose_amount.rds")
i_mort <- readRDS("../fit_models/indiv_mort_dose_amount.rds")
i_sev <- readRDS("../fit_models/indiv_severity_dose_amount.rds")

## Joint plot

padding <- theme(plot.margin = unit(c(t=0.1, r=0.1, b=0, l=0), "cm"))
    
cp <- egg::ggarrange(        
    (intervals_plot(i_dis) + padding + theme(axis.text.y=element_text(size=12)) + ggtitle("A) Disease presence")), 
    (intervals_plot(i_mort) + padding + theme(axis.text.y=element_blank()) + ggtitle("B) Mortality")), 
    (intervals_plot(i_sev) + padding + theme(axis.text.y=element_blank(), legend.position = "right") + ggtitle("C) Severity")), ncol=3, padding=0.5)

ggsave("../plots_tables/combined_individual_intervals_dose_amount.pdf", cp, width=11.5, height=3.5)


```


## Dose and  mass mass as separate predictors

```{r plots_dose_amount_mass, fig.width=10}
  
i_dis <- readRDS("../fit_models/indiv_disease_dose_amount_mass.rds")
i_mort <- readRDS("../fit_models/indiv_mort_dose_amount_mass.rds")
i_sev <- readRDS("../fit_models/indiv_severity_dose_amount_mass.rds")

## Joint plot

padding <- theme(plot.margin = unit(c(t=0.1, r=0.1, b=0, l=0), "cm"))
    
cp <- egg::ggarrange(        
    (intervals_plot(i_dis) + padding + theme(axis.text.y=element_text(size=12)) + ggtitle("A) Disease presence")), 
    (intervals_plot(i_mort) + padding + theme(axis.text.y=element_blank()) + ggtitle("B) Mortality")), 
    (intervals_plot(i_sev) + padding + theme(axis.text.y=element_blank(), legend.position = "right") + ggtitle("C) Severity")), ncol=3, padding=0.5)

ggsave("../plots_tables/combined_individual_intervals_dose_amount_mass.pdf", cp, width=11.5, height=3.5)


```



# Subset to viruses inoculated into both host orders

```{r plots_both_orders, fig.width=10}
  
i_dis <- readRDS("../fit_models/indiv_disease_viruses_both_orders.rds")
i_mort <- readRDS("../fit_models/indiv_mort_viruses_both_orders.rds")
i_sev <- readRDS("../fit_models/indiv_severity_viruses_both_orders.rds")

## Joint plot

padding <- theme(plot.margin = unit(c(t=0.1, r=0.1, b=0, l=0), "cm"))
    
cp <- egg::ggarrange(        
    (intervals_plot(i_dis) + padding + theme(axis.text.y=element_text(size=12)) + ggtitle("A) Disease presence")), 
    (intervals_plot(i_mort) + padding + theme(axis.text.y=element_blank()) + ggtitle("B) Mortality")), 
    (intervals_plot(i_sev) + padding + theme(axis.text.y=element_blank(), legend.position = "right") + ggtitle("C) Severity")), ncol=3, padding=0.5)

ggsave("../plots_tables/combined_individual_intervals_viruses_both_orders.pdf", cp, width=11.5, height=3.5)


```


# Removing inocluations into the brain

```{r plots_no_brain, fig.width=10}
  
i_dis <- readRDS("../fit_models/indiv_disease_no_brain.rds")
i_mort <- readRDS("../fit_models/indiv_mort_no_brain.rds")
i_sev <- readRDS("../fit_models/indiv_severity_no_brain.rds")

## Joint plot

padding <- theme(plot.margin = unit(c(t=0.1, r=0.1, b=0, l=0), "cm"))
    
cp <- egg::ggarrange(        
    (intervals_plot(i_dis) + padding + theme(axis.text.y=element_text(size=12)) + ggtitle("A) Disease presence")), 
    (intervals_plot(i_mort) + padding + theme(axis.text.y=element_blank()) + ggtitle("B) Mortality")), 
    (intervals_plot(i_sev) + padding + theme(axis.text.y=element_blank(), legend.position = "right") + ggtitle("C) Severity")), ncol=3, padding=0.5)

ggsave("../plots_tables/combined_individual_intervals_no_brain.pdf", cp, width=11.5, height=3.5)


```


# Human Case-Fatality Rate models

## Base CFR model

```{r crf_base_models, echo=T, eval=T, message=F, warning=F}

# do bats and rodents differ in mean human CFR of inoculated viruses?
# model with no taxonomic / phylogenetic effects - rodents infected with lower human cfr viruses

cfr_base1 <- readRDS("../fit_models/cfr_base1.rds")

posterior <- as.matrix(cfr_base1)
ar1 <- mcmc_intervals(posterior, regex_pars = "b_[^I]", prob=0.8, prob_outer=0.95, point_est="mean", 
        point_size = 3.5, outer_size=1, inner_size=3) 

# model with only viral taxonomic hierarcical effects

cfr_base2 <- readRDS("../fit_models/cfr_base2.rds")

color_scheme_set("teal")
posterior <- as.matrix(cfr_base2)
ar2 <- mcmc_intervals(posterior, regex_pars = "b_[^I]", prob=0.8, prob_outer=0.95, point_est="mean", 
        point_size = 3.5, outer_size=1, inner_size=3)
# ar2


# model with no viral effects, but removing bat-reservoired viruses

cfr_nobatv <- readRDS("../fit_models/cfr_base3.rds")

posterior <- as.matrix(cfr_nobatv)
ar3 <- mcmc_intervals(posterior, regex_pars = "b_[^I]", prob=0.8, prob_outer=0.95, point_est="mean", 
        point_size = 3.5, outer_size=1, inner_size=3)

# # combo plot
# p1 <- plot_grid(  ar1 + xlim(-3, 2) + scale_y_discrete(labels = "No virus effects\nAll viruses"), 
#                   ar3 + xlim(-3, 2) + scale_y_discrete(labels = "No virus effects\nExcluding bat-reservoired viruses"), 
#                   ar2 + xlim(-3, 2) + scale_y_discrete(labels = "With virus effects\nAll viruses"),
#                   ncol=1, align = "v")
# p2 <- add_sub(p1, "Host order (Chiroptera) effect", x = 0, hjust = -3.5, size=10)
# ggdraw(p2)

# ggsave("../plots_tables/CFR_models_intervals.pdf", width=6, height=4)




### Models with only heterologous hosts

# model with no virus hierarchical effects

cfr_het1 <- readRDS("../fit_models/cfr_het1.rds")

posterior <- as.matrix(cfr_het1)
ar4 <- mcmc_intervals(posterior, regex_pars = "b_[^I]", prob=0.8, prob_outer=0.95, point_est="mean", 
        point_size = 3.5, outer_size=1, inner_size=3)


# model with only viral taxonomic hierarcical effects

cfr_het2 <- readRDS("../fit_models/cfr_het2.rds")

color_scheme_set("teal")
posterior <- as.matrix(cfr_het2)
ar5 <- mcmc_intervals(posterior, regex_pars = "b_[^I]", prob=0.8, prob_outer=0.95, point_est="mean", 
        point_size = 3.5, outer_size=1, inner_size=3)
# ar5

# combo plot
p1 <- plot_grid(  ar1 + xlim(-3, 2) + scale_y_discrete(labels = "No virus effects\nAll viruses"), 
                  ar3 + xlim(-3, 2) + scale_y_discrete(labels = "No virus effects\nExcluding bat-reservoired viruses"), 
                  ar2 + xlim(-3, 2) + scale_y_discrete(labels = "With virus effects\nAll viruses"),
                  ar4 + xlim(-3, 2) + scale_y_discrete(labels = "No virus effects\nHeterologous hosts"),
                  ar5 + xlim(-3, 2) + scale_y_discrete(labels = "With virus effects\nHeterologous hosts") + xlab("Effect of host order (Chiroptera) on human CFR") + theme(axis.title.x = element_text(vjust=-1)),
                  ncol=1, align = "v")
# p2 <- add_sub(p1, "Effect of host order (Chiroptera) on human case-fatality rate", x = 0, hjust = -3.5, size=10)
p1

ggsave("../plots_tables/CFR_models_intervals.pdf", p1, width=6, height=6)


```